When to ignore perfect information: how iterative learning can lead to worse decisions

### 2021-03-26 manuscript/manuscript.Rmd
:pen:

```{r figure5,  fig.cap="Panel A: Population growth curves of each model. The positive equilibrium of each model occurs where the curve crosses the horizontal axis. Note that while Model 2 is a better approximation to the truth (Model 3), Model 1 better approximates the stock size which leads to maximum growth.  Panel B: The optimal control policy under Model 1 is nearly identical to that under the true Model 3, while the optimal policy under Model 2 supresses stock to a much lower escapement level.", fig.width=8}
# Fig 5
model_curves <- read_csv("../data/model_curves.csv") %>% mutate(model = as.character(model))
policies <- read_csv("../data/policies.csv") %>% mutate(model = as.character(model))

@@

<!-- **Is it just a unicorn?** -->
Do models which give nearly optimal performance while at the same time making wildly wrong predictions really exist?  It is tempting to argue that Model 1 is a kind of unicorn, a mythical creature existing only in theory. But that is a difficult assertion to prove.  Examples of serious forecasts that widely fail to predict future observations abound wherever forecasts are common, from elections to economics to environmental change [@].  If there is no shortage of bad forecasts, then could any of them be useful?  The premise that model need not perfectly capture reality to be useful is at the very heart of modeling.  Model 1 successfully captures the one key feature driving decisions in optimal harvest control problems: the stock size at which the maximum growth rate occurs [Fig 5A].  Capturing only the essential aspects as simply as possible is the goal of any model building exercise.  Thus it should be no surprise that a model can drive good decisions while making poor predictions.  It is also important to note that in many cases, any models being used in decision-making are not necessarily being subjected to the rigorous evaluation and updating steps proposed by adaptive management [@Walters1978] or iterative forecasting [@Dietze2018].  This makes it more likely that such models could persist in practice until now. While iterations that revise this models will no doubt improve decision outcomes in many case, it is worth bearing in mind from the example here that such a connection is not guaranteed.

<!-- So what? -->
If these unicorns do exist in real world management, then what do we do about them? I believe this is an open question, but that our first step must be to recognize it as such.  We have seen that the problem cannot be resolved by more data, and is not the result of overfitting. Nor is "creating more models" the answer: when we have a model that is good enough to get optimal results, we cannot always insist on more models. We have also seen how not updating our uncertainty estimates, or doing so less frequently, can reveal a unicorn model before it is discounted by further iterations. Approaches such as iterative forecasting or adaptive management that can reduce model uncertainty over time remain promising and important techniques, but because value of information can be negative, we must revise model uncertainty only cautiously.  More importantly, we have seen that  the intuition offered by decision theory [Fig 5] can help us better understand what features of a model are essential to decision outcomes and what are not.  Conversely, models which appear to lead to accurate predictions (like Model 2) can result in outcomes that are far from optimal. Building on such understanding, it may possible to identify strategies for learning that are more agnostic to the details of the model that are not important, or perhaps not reliant on model-based predictions at all. How that is done is an open question.

>>>>

```{r figure4,  fig.cap="Panel A: Population growth curves of each model. The positive equilibrium of each model occurs where the curve crosses the horizontal axis. Note that while Model 2 is a better approximation to the truth (Model 3), Model 1 better approximates the stock size which leads to maximum growth.  Panel B: The optimal control policy under Model 1 is nearly identical to that under the true Model 3, while the optimal policy under Model 2 supresses stock to a much lower escapement level.", fig.width=8}
# Fig 4
model_curves <- read_csv("../data/model_curves.csv") %>% mutate(model = as.character(model))
policies <- read_csv("../data/policies.csv") %>% mutate(model = as.character(model))

@@

<!-- **Is it just a unicorn?** -->
Do models which give nearly optimal performance while at the same time making wildly wrong predictions really exist?  It is tempting to argue that Model 1 is a kind of unicorn, a mythical creature existing only in theory. But that is a difficult assertion to prove.  Examples of serious forecasts that widely fail to predict future observations abound wherever forecasts are common, from elections to economics to environmental change [e.g. @Tetlock2015].  If there is no shortage of bad forecasts, then could any of them be useful?  The premise that model need not perfectly capture reality to be useful is at the very heart of modeling.  Model 1 successfully captures the one key feature driving decisions in optimal harvest control problems: the stock size at which the maximum growth rate occurs [Fig 4A].  Capturing only the essential aspects as simply as possible is the goal of any model building exercise.  Thus it should be no surprise that a model can drive good decisions while making poor predictions.  It is also important to note that in many cases, any models being used in decision-making are not necessarily being subjected to the rigorous evaluation and updating steps proposed by adaptive management [@Walters1978] or iterative forecasting [@Dietze2018].  This makes it more likely that such models could persist in practice until now. While iterations that revise this models will no doubt improve decision outcomes in many case, it is worth bearing in mind from the example here that such a connection is not guaranteed.

<!-- So what? -->
If these unicorns do exist in real world management, then what do we do about them? I believe this is an open question, but that our first step must be to recognize it as such.  We have seen that the problem cannot be resolved by more data, and is not the result of overfitting. Nor is "creating more models" the answer: when we have a model that is good enough to get optimal results, we cannot always insist on more models. We have also seen how not updating our uncertainty estimates, or doing so less frequently, can reveal a unicorn model before it is discounted by further iterations. Approaches such as iterative forecasting or adaptive management that can reduce model uncertainty over time remain promising and important techniques, but because value of information can be negative, we must revise model uncertainty only cautiously.  More importantly, we have seen that  the intuition offered by decision theory [Fig 4] can help us better understand what features of a model are essential to decision outcomes and what are not.  Conversely, models which appear to lead to accurate predictions (like Model 2) can result in outcomes that are far from optimal. Building on such understanding, it may possible to identify strategies for learning that are more agnostic to the details of the model that are not important, or perhaps not reliant on model-based predictions at all. How that is done is an open question.
